Long-Term Reproducibility through Replication:\\ the ReScience Initiative

### 2017-06-12 draft.tex
Discussion of Software Heritage

>>>>

Based on our experience with the ReScience initiative, we can engage
in informed speculation about possible future evolutions in scientific
publishing, in particular concerning replication work. We won't
discuss minor technical advances such as a better toolchain for
producing PDF articles, but concentrate on long-term improvements in
the technology of electronic publishing and, most of all, in the
attitude of the scientific community towards the publication,
preservation, and verification of computer-aided research.

Today's electronic publishing platforms for scientific research still
show their origins in paper-based publishing. Except for the
replacement of printed paper by a printable PDF file, not much has
changed. Although it is increasingly realized that software and data
should be integral parts of most scientific publications today, they
are at best relegated to the status of ``supplementary material'', and
systematically excluded from the peer review process. In fact, to the
best of our knowledge, ReScience is the only scientific journal that
aims to verify the correctness of scientific software. As our
experience has shown, it is far easier to graft publication onto a
software development platform than to integrate software reviewing
into a publishing platform.

A logical next step is to fully embrace the technology designed for
software development, which far better takes into account the
specificity of electronic information processing than today's
scientific publishing systems. In addition to the proper handling of
code, such an approach offers further advantages. Perhaps the most
important one is a shift of focus from the paper as a mostly isolated
and finished piece of work to scientific progress as a collection of
incremental and highly interdependent steps. The
\href{https://www.softwareheritage.org/}{Software Heritage} project,
whose aim is to create a permanent public archive of all publicly
available software source code, adopts exactly this point of view for
the preservation of software. As our experience with ReScience has
shown, integrating the narrative of a scientific article into a
framework designed for software development is not difficult at all.
Publishing and archiving scientific research in Software Heritage
would offer several advantages. The intrinsic identifiers that provide
access to the contents of the archive permit unambiguous and permanent
references to ongoing projects as well as to snapshots at a specific
time, and to whole projects as well as to the individual files that
are part of them. Such references hold the promise for better reuse of
scientific information, for better reproducibility of
computations, and for fairer attribution of credit to scientists
who contribute to research infrastructure.
